Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-07-14 main.tex
partially section 5

The main objective of this manuscript is bounding the Poincar\'e--Friedrichs constants over local stars within triangulated domains. 
Clearly, not all local stars describe convex subdomains. 
One possible avenue towards Poincar\'e--Friedrichs constants is the observation that local stars in triangulations are star-shaped with respect to a ball,
which enables tools such as averaged Poincar\'e and Bogovski\u{\i} operators. 
However, estimates that rely on this geometric condition deteriorate when the aforementioned ball has radius much smaller than the domain diameter. 
This is not as much a problem over patches around interior subsimplices, where the size of the interior ball only depends on the shape regularity of the triangulation. 
But that interior ball can be arbitrarily small when the patch is around a boundary simplex, even if the mesh has good shape regularity: 
this occurs when the domain has sharp reentrant corners, and some illustrative limit cases include the slit domain~\cite{veeser2012poincare} and the crossed bricks domain~\cite{licht2019smoothed}, % DONE: MV : find the examples in the paper by verfuhrt and veeser 
which contain finite element patches are not star-shaped with respect to any ball.

@@

When $\calT$ is a triangulation and $T \in \calT$, then $\patch_{\calT}(T)$ denotes the \emph{local patch} or \textit{local star} of $T$, 
which the simplicial subcomplex of $\calT$ that contains all supersimplices of $T$ and their subsimplices. Formally,
\begin{gather*}
    \patch_{\calT}(T) := \bigcup_{ T' \in \supsimplex(\calT,T) } \subsimplex(T').
\end{gather*}
We write $U_T := | \patch_{\calT}(T) |$ for the underlying set of the local patch. 
A crucial observation is the following.

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulated manifolds. 
We put particular focus on \emph{shellable} simplicial complexes as a particularly important class of complexes. Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. Patches within triangulations of dimension two and three are examples of such shellable complexes. 
Our main references for this section are the monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures}.

@@

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its midpoint. 
    Let $\mathring B_F$ be an open neighborhood around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, and their interiors are disjoint.

@@

\begin{lemma}\label{lemma:startopology}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item If $S$ is an inner simplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. % TODO: Define this macro 
        \item If $S$ is a boundary subsimplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
        \item If the underlying space of $\calT$ is connected, then $\calT$ is strongly connected.
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove these statements in several steps.
    \begin{enumerate}
    \item 
    Let $S$ be any simplex with vertices $v_0, v_1, \dots, v_k$, with midpoint $z_F$, and dimension $k$.
    Let $\calS := \patch_{\calT}(S)$ be its star. 
    Each $l$-dimensional simplex $T \in \calS$ that contains $S$ 
    has vertices $v_0, v_1, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}$. 
    For any such simplex, we introduce a decomposition $T_{0}, \dots, T_{k}$, where each $T_{i}$ has vertices 
    \begin{align*}
        v_0, \dots, v_{i-1}, z_F, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}.
    \end{align*}
    The collection $\calS'$ of these simplices and their subsimplices constitute a simplicial complex 
    that triangulates the same underlying set as $\calS$.
    Moreover, $\calS' = \patch_{\calS'}(z_F)$. 
    In particular, $z_F$ is a boundary vertex of $\calS'$ if and only if $S$ is a boundary simplex of $\calS$.

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- cycle;
        \draw[dashed] (O) -- (A); \draw[dotted] (O) -- (B); \draw (O) -- (C); \draw (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}
    \begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw[dotted] (A) -- (B); \draw (B) -- (C) -- (D) -- (E) -- (F) -- (A);
        \draw[dashed] (O) -- (A); \draw (O) -- (B); \draw (O) -- (C); \draw (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}
    \begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- cycle;
        \draw[dashed] (O) -- (A); \draw (O) -- (B); \draw[dotted] (O) -- (C); \draw (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}
    \begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- cycle;
        \draw[dashed] (O) -- (A); \draw (O) -- (B); \draw (O) -- (C); \draw[dotted] (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}
    \caption{Examples and counterexamples for building a manifold-like $1$-patching of a $2$-dimensional manifold-like triangulation. (1) We have already added the edge patch along the dashed line and we add the edge patch along the dotted line. (2) The new edge patch may already be a subset of the subtriangulation (3) An inadmissible choice because new edge patch has no overlap with current subtriangulation, though the next subtriangulation would be manifold-like (4) An inadmissible choice because new edge patch has no overlap with current subtriangulation and the next subtriangulation would not be manifold-like.}\label{figure:illustrationpatching}
\end{figure}

\begin{figure}
    \centering
    \begin{tikzpicture}[scale=0.5]
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({0.8*cos(  0)}, {0.8*sin(  0)});
        \coordinate (B) at ({0.8*cos(120)}, {0.8*sin(120)});
        \coordinate (C) at ({0.8*cos(240)}, {0.8*sin(240)});
        \coordinate (X) at ({3.0*cos( 60)}, {3.0*sin( 60)});
        \coordinate (Y) at ({3.0*cos(180)}, {3.0*sin(180)});
        \coordinate (Z) at ({3.0*cos(300)}, {3.0*sin(300)});
        \draw (A) -- (B) -- (C) -- cycle; \draw (X) -- (Y) -- (Z) -- cycle;
        \draw[fill=gray!20] (X) -- (Y) -- (B) -- cycle;
        \draw[fill=gray!20] (Y) -- (Z) -- (C) -- cycle;
        \draw[fill=gray!20] (Z) -- (X) -- (A) -- cycle;
        \draw[fill=gray!20] (A) -- (X) -- (B) -- cycle;
        \draw[fill=gray!20] (B) -- (Y) -- (C) -- cycle;
        \draw[fill=gray!20] (C) -- (Z) -- (A) -- cycle;
    \end{tikzpicture}
    \begin{tikzpicture}[scale=0.5]
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({3.0*cos(  0)}, {3.0*sin(  0)});
        \coordinate (B) at ({3.0*cos(120)}, {3.0*sin(120)});
        \coordinate (C) at ({3.0*cos(240)}, {3.0*sin(240)});
        \coordinate (X) at ({3.0*cos( 60)}, {3.0*sin( 60)});
        \coordinate (Y) at ({3.0*cos(180)}, {3.0*sin(180)});
        \coordinate (Z) at ({3.0*cos(300)}, {3.0*sin(300)});
        \draw (O) -- (A);
        \draw (O) -- (B);
        \draw (O) -- (C);
        \draw (O) -- (X);
        \draw (O) -- (Y);
        \draw (O) -- (Z);
        %
    \end{tikzpicture}
    \begin{tikzpicture}[scale=0.5] 
        \tikzset{z={(90:1cm)}, y={(-30:1cm)}, x={(210:1cm)}}
        % \tikzset{z={(90:1cm)}, y={(-10:1cm)}, x={(250:1cm)}}

        \coordinate (O) at (0, 0);
        \coordinate (A) at ({3.0*cos(  0)}, {3.0*sin(  0)});
        \coordinate (B) at ({3.0*cos(120)}, {3.0*sin(120)});
        \coordinate (C) at ({3.0*cos(240)}, {3.0*sin(240)});
        \coordinate (X) at ({3.0*cos( 60)}, {3.0*sin( 60)});
        \coordinate (Y) at ({3.0*cos(180)}, {3.0*sin(180)});
        \coordinate (Z) at ({3.0*cos(300)}, {3.0*sin(300)});
        \coordinate (T) at ({0}, {0}, {3});

        \draw[thick, fill=gray!30] (O) -- (C) -- (T) -- cycle;
        \draw[thick, fill=gray!30] (O) -- (Y) -- (T) -- cycle;
        \draw[thick, fill=gray!30] (O) -- (Z) -- (T) -- cycle;
        \draw[thick, fill=gray!30] (O) -- (A) -- (T) -- cycle;
        \draw[thick, fill=gray!30] (O) -- (B) -- (T) -- cycle;
        \draw[thick, fill=gray!30] (O) -- (X) -- (T) -- cycle;

    \end{tikzpicture}

    \caption{Example of a connected $2$-dimensional manifold triangulation, which triangulates an annulus, that does neither admit a manifold-like $0$-patching nor manifold-like $1$-patching.}\label{figure:annuluscounterexample}
\end{figure}

@@

\begin{remark}
    We interpret a shelling as the construction of a triangulation 
    by successively attaching simplices such that the intermediate triangulations are well-behaved. 
    Conversely, the reverse enumeration describes a successive decomposition of the triangulation, hence the name ``shelling''.
\end{remark}
\begin{remark}
    Whether a simplicial complex is shellable can be checked, in principle, simply by trying out all the possibly enumerations.
    That we cannot do much better than this is captured in the result that testing for shellability is NP-complete~\cite{goaoc2019shellability}:
    this complexity result is even true if we merely consider simplicial complexes of dimension two embedded in some Euclidean space.
\end{remark}

@@

\begin{lemma}    
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, % TODO: what is non-branching?
    then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

@@

    Write $M = |\calT|$. 
    We call a triangle $T \in \calT$ \emph{good} if it intersects the boundary $\partial M$ in a non-empty union of edges. % TODO: I understand what it means from the rest of the proof, but I believe it could a bit more explicit.
    We show by an induction argument over the number of triangles that every triangulation of a $2$-disk contains a good triangle. 
    Clearly, this is the case if the triangulation contains only one triangle. 
    Now suppose the induction claim is true when the triangulation includes $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
    There exists a triangle $T' \in \calT$ with at least one edge in the boundary. 
    If $T'$ is good, then we are already done. 
    Otherwise, $T'$ intersects the boundary at one edge and its opposing vertex.
    Removing $T'$ splits the manifold into two pieces, each of which is a topological $2$-disk.
    By the induction assumption, they contain a good triangle.

    Since $\calT$ triangulates a $2$-disk,
    it contains a good triangle $T$. 
    If $T$ has three edges in the boundary, then $T = M$ and we are trivially done. 
    If $T$ intersects with the boundary in exactly one or two edges, 
    then $\overline{M \setminus T}$ is still a topological $2$-disk.
    Letting $\calT'$ be the triangulation obtained by removing $T$,
    we have another triangulation of the $2$-disk that intersects $T$ only at either two or one edges.
    Any shelling of $\calT'$ can be extended to a shelling of $\calT$.
%     Let $\calT$ be a simplicial $2$-disk. 
%     If $\calT$ contains only one triangle, then there is nothing to show. 
%     Otherwise, there exists at least one triangle $T$ that touches the boundary of the disk at either one or two edges 
%     {\color{red}Needs proof}. 
%   
%     Now let $\calS$ be any simplicial $2$-sphere. 
%     By removing any triangle $S \in \calS$, we obtain a simplicial $2$-disk, which is shellable. 
%     Any of its shellings can be extended to a shelling of $\calS$ by re-inserting the first triangle $S$.
\end{proof}

\begin{lemma}
    Let $\calT$ be a $n$-dimensional manifold triangulation and $V \in \Vertices(\calT)$ be an inner vertex.
    Then $\patch_{\calT}(V)$ is shellable. 
\end{lemma}
\begin{proof}
    Without loss of generality, $V$ is the origin. 
    Any ray starting at $V$ intersects at most one vertex of $\patch_{\calT}(V)$.
    We can also assume that all boundary vertices of $\patch_{\calT}(V)$ to be unit vectors,
    which does not affect shellability. 
    The simplicial boundary complex of $\calT$ is the boundary complex of a convex polytope,
    and is thus shellable~\cite[Theorem~8.12]{ziegler2012lectures}. 
\end{proof}

>>>>

The main objective of this manuscript is bounding the Poincar\'e--Friedrichs constants over local stars within triangulated domains. 
Clearly, not all local stars describe convex subdomains. 
One possible avenue towards Poincar\'e--Friedrichs constants is the observation that local stars in triangulations are star-shaped with respect to a ball,
which enables tools such as averaged Poincar\'e and Bogovski\u{\i} operators. 
However, estimates that rely on this geometric condition deteriorate when the aforementioned ball has radius much smaller than the domain diameter. 
This is not as much a problem over patches around interior subsimplices, where the size of the interior ball only depends on the shape regularity of the triangulation. 
But that interior ball can be arbitrarily small when the patch is around a boundary simplex, even if the mesh has good shape regularity: 
this occurs when the domain has sharp reentrant corners, and some illustrative limit cases include the slit domain~\cite{veeser2012poincare} and the crossed bricks domain~\cite{licht2019smoothed}, % DONE: MV : find the examples in the paper by verfurth and veeser 
which contain finite element patches are not star-shaped with respect to any ball.

@@

When $\calT$ is a triangulation and $T \in \calT$, then $\patch_{\calT}(T)$ denotes the \emph{local patch} or \textit{local star} of $T$, 
which the simplicial subcomplex of $\calT$ that contains all supersimplices of $T$ and their subsimplices. 
We write $\carapace_{\calT}(T)$ for the subset of the local patch whose members do not contain $T$ itself. 
Formally,
\begin{gather*}
    \patch_{\calT}(T) := \bigcup_{ T' \in \supsimplex(\calT,T) } \subsimplex(T'),
    \qquad 
    \carapace_{\calT}(T) := \bigcup_{ \substack{ T' \in \patch_{\calT}(T) \\ T \nsubseteq T' } } \subsimplex(T').
\end{gather*}
We write $U_T := | \patch_{\calT}(T) |$ for the underlying set of the local patch. 
A crucial observation is the following.

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulated manifolds. 
We put particular focus on \emph{shellable} simplicial complexes as an important class of complexes. Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. Patches within triangulations of dimension two and three are examples of such shellable complexes. 
Our main references for this section are the monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler2012lectures}.

@@

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its barycenter. 
    Let $\mathring B_F$ be an open neighborhood around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, and their interiors are disjoint.

@@

\begin{lemma}\label{lemma:startopology}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$. 
    \begin{itemize}
        \item If $S$ is an inner simplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as an inner simplex
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-sphere. 
        \item If $S$ is a boundary subsimplex, then $\patch_{\calT}(S)$ is a simplicial $n$-ball with $S$ as a boundary simplex,
        and $\carapace_{\calT}(S)$ is a simplicial $(n-1)$-ball.
        \item If the underlying space of $\calT$ is connected, then $\calT$ is strongly connected.
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove these statements in several steps.
    \begin{enumerate}
    \item 
    Let $S$ be any simplex with vertices $v_0, v_1, \dots, v_k$, with midpoint $z_F$, and dimension $k$.
    Let $\calS := \patch_{\calT}(S)$ be its star. 
    Each $l$-dimensional simplex $T \in \calS$ that contains $S$ 
    has vertices $v_0, v_1, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}$. 
    For any such simplex, we introduce a decomposition $T_{0}, \dots, T_{k}$, where each $T_{i}$ has vertices 
    \begin{align*}
        v_0, \dots, v_{i-1}, z_F, v_{i+1}, \dots, v_k, v_{k+1}^{S}, \dots, v_{l}^{S}.
    \end{align*}
    The collection $\calS'$ of these simplices and their subsimplices constitute a simplicial complex 
    that triangulates the same underlying set as $\calS$.
    Moreover, $\calS' = \patch_{\calS'}(z_F)$. 
    In particular, $z_F$ is a boundary vertex of $\calS'$ if and only if $S$ is a boundary simplex of $\calS$.

@@

\begin{figure}
    \centering
    \begin{tabular}{c}\begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- cycle;
        \draw[dashed] (O) -- (A); \draw[dotted] (O) -- (B); \draw (O) -- (C); \draw (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}\\ (1) \end{tabular}
    \begin{tabular}{c}\begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw[dotted] (A) -- (B); \draw (B) -- (C) -- (D) -- (E) -- (F) -- (A);
        \draw[dashed] (O) -- (A); \draw (O) -- (B); \draw (O) -- (C); \draw (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}\\ (2) \end{tabular}
    \begin{tabular}{c}\begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- cycle;
        \draw[dashed] (O) -- (A); \draw (O) -- (B); \draw[dotted] (O) -- (C); \draw (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}\\ (3) \end{tabular}
    \begin{tabular}{c}\begin{tikzpicture}
        \coordinate (O) at (0, 0);
        \coordinate (A) at ({cos(0)}, {sin(0)});
        \coordinate (B) at ({cos(60)}, {sin(60)});
        \coordinate (C) at ({cos(120)}, {sin(120)});
        \coordinate (D) at ({cos(180)}, {sin(180)});
        \coordinate (E) at ({cos(240)}, {sin(240)});
        \coordinate (F) at ({cos(300)}, {sin(300)});
        \fill[gray!20] (A) -- (B) -- (O) -- (F) -- cycle;
        \draw (A) -- (B) -- (C) -- (D) -- (E) -- (F) -- cycle;
        \draw[dashed] (O) -- (A); \draw (O) -- (B); \draw (O) -- (C); \draw[dotted] (O) -- (D); \draw (O) -- (E); \draw (O) -- (F);
    \end{tikzpicture}\\ (4) \end{tabular}
    \caption{Examples and counterexamples for building a manifold-like $1$-patching of a $2$-dimensional manifold-like triangulation. 
    (1) We have already added the edge patch along the dashed line and we add the edge patch along the dotted line.
    (2) The new edge patch may already be a subset of the subtriangulation
    (3) An inadmissible choice because the new edge patch has no overlap with current subtriangulation, though the next subtriangulation would be manifold-like
    (4) An inadmissible choice because the new edge patch has no overlap with current subtriangulation and the next subtriangulation would not be manifold-like.}\label{figure:illustrationpatching}
\end{figure}

% \begin{figure}
%     \centering
%     \begin{tabular}{c}\begin{tikzpicture}[scale=0.5]
%         \coordinate (O) at (0, 0);
%         \coordinate (A) at ({0.8*cos(  0)}, {0.8*sin(  0)});
%         \coordinate (B) at ({0.8*cos(120)}, {0.8*sin(120)});
%         \coordinate (C) at ({0.8*cos(240)}, {0.8*sin(240)});
%         \coordinate (X) at ({3.0*cos( 60)}, {3.0*sin( 60)});
%         \coordinate (Y) at ({3.0*cos(180)}, {3.0*sin(180)});
%         \coordinate (Z) at ({3.0*cos(300)}, {3.0*sin(300)});
%         \draw (A) -- (B) -- (C) -- cycle; \draw (X) -- (Y) -- (Z) -- cycle;
%         \draw[fill=gray!20] (X) -- (Y) -- (B) -- cycle;
%         \draw[fill=gray!20] (Y) -- (Z) -- (C) -- cycle;
%         \draw[fill=gray!20] (Z) -- (X) -- (A) -- cycle;
%         \draw[fill=gray!20] (A) -- (X) -- (B) -- cycle;
%         \draw[fill=gray!20] (B) -- (Y) -- (C) -- cycle;
%         \draw[fill=gray!20] (C) -- (Z) -- (A) -- cycle;
%     \end{tikzpicture}\\ (1) \end{tabular}
%     \begin{tabular}{c}\begin{tikzpicture}[scale=0.5]
%         \coordinate (O) at (0, 0);
%         \coordinate (A) at ({3.0*cos(  0)}, {3.0*sin(  0)});
%         \coordinate (B) at ({3.0*cos(120)}, {3.0*sin(120)});
%         \coordinate (C) at ({3.0*cos(240)}, {3.0*sin(240)});
%         \coordinate (X) at ({3.0*cos( 60)}, {3.0*sin( 60)});
%         \coordinate (Y) at ({3.0*cos(180)}, {3.0*sin(180)});
%         \coordinate (Z) at ({3.0*cos(300)}, {3.0*sin(300)});
%         \draw (O) -- (A);
%         \draw (O) -- (B);
%         \draw (O) -- (C);
%         \draw (O) -- (X);
%         \draw (O) -- (Y);
%         \draw (O) -- (Z);
%         %
%     \end{tikzpicture}\\ (2) \end{tabular}
%     \begin{tabular}{c}\begin{tikzpicture}[scale=0.5]
%         \tikzset{z={(90:1cm)}, y={(-30:1cm)}, x={(210:1cm)}}
%         % \tikzset{z={(90:1cm)}, y={(-10:1cm)}, x={(250:1cm)}}
% 
%         \coordinate (O) at (0, 0);
%         \coordinate (A) at ({3.0*cos(  0)}, {3.0*sin(  0)});
%         \coordinate (B) at ({3.0*cos(120)}, {3.0*sin(120)});
%         \coordinate (C) at ({3.0*cos(240)}, {3.0*sin(240)});
%         \coordinate (X) at ({3.0*cos( 60)}, {3.0*sin( 60)});
%         \coordinate (Y) at ({3.0*cos(180)}, {3.0*sin(180)});
%         \coordinate (Z) at ({3.0*cos(300)}, {3.0*sin(300)});
%         \coordinate (T) at ({0}, {0}, {3});
% 
%         \draw[thick, fill=gray!30] (O) -- (C) -- (T) -- cycle;
%         \draw[thick, fill=gray!30] (O) -- (Y) -- (T) -- cycle;
%         \draw[thick, fill=gray!30] (O) -- (Z) -- (T) -- cycle;
%         \draw[thick, fill=gray!30] (O) -- (A) -- (T) -- cycle;
%         \draw[thick, fill=gray!30] (O) -- (B) -- (T) -- cycle;
%         \draw[thick, fill=gray!30] (O) -- (X) -- (T) -- cycle;
% 
%     \end{tikzpicture}\\ (3) \end{tabular}
%     \caption{Example of a connected $2$-dimensional manifold triangulation, which triangulates an annulus, that does neither admit a manifold-like $0$-patching nor manifold-like $1$-patching.}\label{figure:annuluscounterexample}
% \end{figure}

@@

\begin{remark}
    We interpret a shelling as the construction of a triangulation 
    by successively attaching simplices such that the intermediate triangulations are well-behaved. 
    Conversely, the reverse enumeration describes a successive decomposition of the triangulation, hence the name ``shelling''.
\end{remark}
\begin{remark}
    Whether a simplicial complex is shellable can be checked, in principle, simply by trying out all the possible enumerations.
    That we cannot do much better than this is captured in the result that testing for shellability is NP-complete~\cite{goaoc2019shellability}:
    this complexity result is even true if we merely consider simplicial complexes of dimension two embedded in some Euclidean space.
\end{remark}

@@

\begin{lemma}    
    Suppose that an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, 
    that is, each simplex of dimension $n-1$ is contained in at most two simplices. 
    % TODO: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

@@

    \todo{Double check this proof and make it more explicit.}
    Write $M = |\calT|$. 
    We call a triangle $T \in \calT$ \emph{good} if it intersects the boundary $\partial M$ in a non-empty union of edges. % TODO: I understand what it means from the rest of the proof, but I believe it could a bit more explicit.
    We show by an induction argument over the number of triangles that every triangulation of a $2$-disk contains a good triangle. 
    Clearly, this is the case if the triangulation contains only one triangle. 
    Now suppose the induction claim is true when the triangulation includes $N$ triangles,
    and assume that $\calT$ includes $N+1$ triangles. 
    There exists a triangle $T' \in \calT$ with at least one edge in the boundary. 
    If $T'$ is good, then we are already done. 
    Otherwise, $T'$ intersects the boundary at one edge and its opposing vertex.
    Removing $T'$ splits the manifold into two pieces, each of which is a topological $2$-disk.
    By the induction assumption, they contain a good triangle.
    \todo{MV:But what if this is the neighbor of $T'$ and adding $T'$ back makes the triangle bad?}

    Since $\calT$ triangulates a $2$-disk,
    it contains a good triangle $T$. 
    If $T$ has three edges in the boundary, then $T = M$ and we are trivially done. 
    If $T$ intersects with the boundary in exactly one or two edges, 
    then $\overline{M \setminus T}$ is still a topological $2$-disk.
    The triangulation $\calT'$ that is obtained by removing $T$
    is a triangulation of some $2$-disk that intersects $T$ only at either two or one edges.
    Any shelling of $\calT'$ can be extended to a shelling of $\calT$.
%     Let $\calT$ be a simplicial $2$-disk. 
%     If $\calT$ contains only one triangle, then there is nothing to show. 
%     Otherwise, there exists at least one triangle $T$ that touches the boundary of the disk at either one or two edges 
%     {\color{red}Needs proof}. 
%   
%     Now let $\calS$ be any simplicial $2$-sphere. 
%     By removing any triangle $S \in \calS$, we obtain a simplicial $2$-disk, which is shellable. 
%     Any of its shellings can be extended to a shelling of $\calS$ by re-inserting the first triangle $S$.
\end{proof}

\begin{lemma}
    Let $\calT$ be a $n$-dimensional manifold triangulation and $V \in \Vertices(\calT)$ be an inner vertex.
    Then $\patch_{\calT}(V)$ is shellable. 
\end{lemma}
\begin{proof}
    Without loss of generality, $V$ is the origin. 
    Any ray starting at $V$ intersects at most one vertex of $\patch_{\calT}(V)$.
    We can also assume that all boundary vertices of $\patch_{\calT}(V)$ to be unit vectors,
    which does not affect shellability. 
    The simplicial boundary complex of $\calT$ is the boundary complex of a convex polytope,
    and is thus shellable~\cite[Theorem~8.12]{ziegler2012lectures}. 
    % TODO: cite E&V Math Comp 2020, Lemma B1. 
\end{proof}
